Diverging taxonomic and functional trajectories following disturbance in a Neotropical forest

### 2020-03-30 WholePlotTrajectories.Rmd
cleanup

---
title: "Diverging taxonomic and functional trajectories following disturbance in a Neotropical forest"
author:
  - name: Ariane MIRABEL
    authsuperscript: 1*
  - name: Bruno Herault
    authsuperscript: 2
  - name: Eric Marcon
    authsuperscript: 1
affiliation:
  - affsuperscript: 1
    dptuniv: UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles, Université de Guyane.
    address: >
      Campus Agronomique,
      97310 Kourou,
      France.
  - affsuperscript: 2
    dptuniv: INPHB, Institut National Polytechnique Félix Houphoüet-Boigny
    address: >
      Yamoussoukro,
      Ivory Coast.
corrauthor:
    email: ariane.mirabel@gmail.com
    url: https://github.com/ArianeMirabel
abstract: 
 In the current global change context, it is urgent to anticipate the fate of tropical forests.
 This means understanding tree community response to disturbance and the underlying processes.
 In that respect, we aim here to clarify taxonomic and functional post-disturbance  trajectories, and determine the scope of the Intermediate Disturbance Hypothesis (IDH) that remains debated in tropical forests.
 \color{black}We analyzed community trajectories following a disturbance gradient from 10 to 60\% of above-ground biomass loss in a Neotropical forest over 30 years.\color{red} 
 We considered trajectories along time of community taxonomic and functional trajectories in terms of richness, evenness, composition, and redundancy.
 We based on the annual botanical inventories of 75 ha of a Neotropical forest and on large trait datasets comprising seven leaf, stem, and life-history traits. 
 We identified a decoupling between taxonomic composition, differing among communities, and functional composition, \color{red} similar among communities and convergent in the functional space.\color{black} 
 \color{red} The taxonomic diversity followed humped-shaped trajectories along time after disturbance depending on the initial disturbance intensity, which validated the IDH (Intermediate Disturbance Hypothesis).\color{black} 
 The functional diversity trajectories,however, were homogeneous among plots and dismissed the IDH.
 We explained this decoupling by the variations in community functional redundancy that mitigated the functional impact of disturbance. 
 Although consistent, the recovery of community composition, diversity, and redundancy remained unachieved after 30 years. \color{red} These results acknowledged the need of decades-long cycles without disturbance to ensure community complete recovery.\color{black} 
keywords: [Community Ecology, Disturbance Trajectories, Intermediate Disturbance Hypothesis, Mid-term Resilience, Neotropical Forests, Taxonomic and Functional Biodiversity]
preamble: >
  \usepackage{textcomp}
  \usepackage{tabu}
  \usepackage{caption}
  \captionsetup{justification = justified}
  \renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
  \hyphenation{bio-di-ver-si-ty sap-lings re-or-gan-i-za-tion post-dis-tur-bance dis-tur-bance}
  \usepackage[left]{lineno}
  \linenumbers
  \usepackage{bookmark}
bibliography: references.bib
lang: english
keywordlabel: Keywords
corrauthorlabel: Corresponding author
output:
  bookdown::gitbook: default
  bookdown::pdf_book:
    base_format: EcoFoG::article
    keep_tex: TRUE
  bookdown::word_document2: default
---

>>>>

---
title: "Diverging taxonomic and functional trajectories following disturbance in a Neotropical forest"
author:
  - name: Ariane MIRABEL
    authsuperscript: 1*
  - name: Bruno Herault
    authsuperscript: 2
  - name: Eric Marcon
    authsuperscript: 1
affiliation:
  - affsuperscript: 1
    dptuniv: UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles, Université de Guyane.
    address: >
      Campus Agronomique,
      97310 Kourou,
      France.
  - affsuperscript: 2
    dptuniv: INPHB, Institut National Polytechnique Félix Houphoüet-Boigny
    address: >
      Yamoussoukro,
      Ivory Coast.
corrauthor:
    email: ariane.mirabel@gmail.com
    url: https://github.com/ArianeMirabel
abstract: 
 In the current global change context, it is urgent to anticipate the fate of tropical forests.
 This means understanding tree community response to disturbance and the underlying processes.
 In that respect, we aim here to clarify taxonomic and functional post-disturbance  trajectories, and determine the scope of the Intermediate Disturbance Hypothesis (IDH) that remains debated in tropical forests.
 \color{black}We analyzed community trajectories following a disturbance gradient from 10 to 60\% of above-ground biomass loss in a Neotropical forest over 30 years.\color{red} 
 We considered trajectories along time of community taxonomic and functional trajectories in terms of richness, evenness, composition, and redundancy.
 We based on the annual botanical inventories of 75 ha of a Neotropical forest and on large trait datasets comprising seven leaf, stem, and life-history traits. 
 We identified a decoupling between taxonomic composition, differing among communities, and functional composition, \color{red} similar among communities and convergent in the functional space.\color{black} 
 \color{red} The taxonomic diversity followed humped-shaped trajectories along time after disturbance depending on the initial disturbance intensity, which validated the IDH (Intermediate Disturbance Hypothesis).\color{black} 
 The functional diversity trajectories,however, were homogeneous among plots and dismissed the IDH.
 We explained this decoupling by the variations in community functional redundancy that mitigated the functional impact of disturbance. 
 Although consistent, the recovery of community composition, diversity, and redundancy remained unachieved after 30 years. \color{red} These results acknowledged the need of decades-long cycles without disturbance to ensure community complete recovery.\color{black} 
keywords: [Community Ecology, Disturbance Trajectories, Intermediate Disturbance Hypothesis, Mid-term Resilience, Neotropical Forests, Taxonomic and Functional Biodiversity]
preamble: >
  \usepackage{textcomp}
  \usepackage{tabu}
  \usepackage{caption}
  \captionsetup{justification = justified}
  \renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
  \hyphenation{bio-di-ver-si-ty sap-lings re-or-gan-i-za-tion post-dis-tur-bance dis-tur-bance}
  \usepackage[left]{lineno}
  \linenumbers
  \usepackage{bookmark}
bibliography: references.bib
lang: english
keywordlabel: Keywords
corrauthorlabel: Corresponding author
output:
  bookdown::pdf_book:
    base_format: EcoFoG::article
    keep_tex: TRUE
  bookdown::gitbook: default
  bookdown::word_document2: default
---

### 2020-03-30 WholePlotTrajectories.tex
cleanup

% User-adder preamble
\usepackage{textcomp} \usepackage{tabu} \usepackage{caption}
\captionsetup{justification = justified}
\renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
\hyphenation{bio-di-ver-si-ty sap-lings re-or-gan-i-za-tion post-dis-tur-bance dis-tur-bance}
\usepackage[left]{lineno} \linenumbers \usepackage{bookmark}

@@

\section{Introduction}\label{introduction}

@@

\section{Material and Methods}\label{material-and-methods}

\subsection{Study site}\label{study-site}

@@

\begin{table}

\caption{\label{tab:Tab1}Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou. Treatment intensities are defined by the minimum logging DBH (Diameter at Breast Height), the type of logged species (commercially exploited or not), the density of logged trees, and the total AGB (Above Ground Biomass) lost after treatment.}
\centering
\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\toprule
Treatment & Timber & Thinning & Fuelwood & \%AGB lost\\
\midrule
Control & - & - & - & 0\\
T1, low & DBH $\geq$ 50 cm, commercially exploited species, $\approx$ 10   $trees.ha^{-1}$ & - & - & $[12-33]$\\
T2, intermediate & DBH $\geq$ 50 cm, commercially exploited species, $\approx$ 10  $trees.ha^{-1}$ & DBH $\geq$ 40 cm, non-commercially exploited species, $\approx$ 30   $trees.ha^{-1}$ & - & $[33-56]$\\
T3, high & DBH $\geq$ 50 cm, commercially exploited species, $\approx$ 10  $trees.ha^{-1}$ & DBH $\geq$ 50 cm, non-commercially exploited species, $\approx$ 15  $trees.ha^{-1}$ & 40 cm $\leq$ DBH $\leq$ 50 cm, non-commercially exploited species,\ $\approx$ 15 $trees.ha^{-1}$ & $[35-56]$\\
\bottomrule
\end{tabu}
\end{table}

\subsection{Inventories protocol and dataset
collection}\label{inventories-protocol-and-dataset-collection}

@@

\subsection{Composition and diversity
metrics}\label{composition-and-diversity-metrics}

@@

\section{Results}\label{results}

\subsection{Community Composition}\label{community-composition}

@@

\subsection{Community taxonomic and functional
diversity}\label{community-taxonomic-and-functional-diversity}

@@

\subsection{Functional redundancy}\label{functional-redundancy}

@@

\section{Discussion}\label{discussion}

@@

\subsection{Decoupled taxonomic and functional
trajectories}\label{decoupled-taxonomic-and-functional-trajectories}

@@

\color{red} We must temper these results with the low number of plots
(3) for each treatment. However, the size of each plot is 6.25ha that
has been regularly censused over 25 years so that the trajectories have
been drawn using 75ha of permanent plot information, a spatial hold very
rarely achieved in such studies. Given the relative homogeneity of
environmental characteristics of terra firme forests over the Guiana
Shield and even the amazonian basin, it seems reasonable that similar
trajectories would apply to neighboring areas
\citep{Guitet2015, Guitet2018}. \color{black}

\subsection{The scope of the intermediate disturbance
hypothesis}\label{the-scope-of-the-intermediate-disturbance-hypothesis}

Trajectories of taxonomic richness and evenness differed markedly below
and above an intensity threshold (Fig. \ref{fig:DivTaxo}). For low and
intermediate disturbance, both taxonomic richness and evenness increased
according to the disturbance intensity
\citep{Martin2015, Chaudhary2016}. This suggested that the recruitment
of pioneers, previously infrequent or absent, increased the taxonomic
richness, and that trees surviving after disturbance remained numerous
enough to maintain the richness of the pre-disturbance community
\citep{Bongers2009}. The pioneers thus recruited became abundant, which
balanced the usual hyper-dominance of tropical forests and increased the
taxonomic evenness \citep{Baraloto2012a}. Above the intensity threshold,
like for treatment 3 (high intensity), the taxonomic richness did not
exceed the initial value in the first years following disturbance. No
increase of the taxonomic richness suggested that the richness of
surviving trees was lower than this of the pre-disturbance community,
and that this difference was not offset by the recruitment of pioneers.
In the Guiana shield, the pool of \color{red}hard \color{black} pioneers
specifically recruited after disturbance is restricted to a few common
genera (e.g. \emph{Cecropia spp.}, \emph{Miconia spp.}, \emph{Tapirira
spp.}) \citep{Guitet2018}. \color{red}The disturbance intensity showed a
humped-shaped relationship with the post-disturbance increase in
taxonomic richness at all the different times after disturbance, and
with the increase in taxonomic evenness 10 years after disturbance (Fig.
\ref{fig:IDHplot}).\color{black} Taxonomic richness, and to some extend
taxonomic evenness, were maximized at an intermediate intensity, around
20-25\% of AGB lost.

\subsection{The functional redundancy explaining the
taxonomic-functional
decoupling}\label{the-functional-redundancy-explaining-the-taxonomic-functional-decoupling}

@@

\section{Conclusion}\label{conclusion}

@@

\section{Acknowledgement}\label{acknowledgement}

@@

\section{Author's contributions}\label{authors-contributions}

@@

\section{Data availability}\label{data-availability}

@@

\bibliographystyle{mee}
\makeatletter
% The filename has .bib extension that must be eliminated
\filename@parse{references.bib}
% parse stores the file name in base. Extension starts at the first dot, so don't use dots in file names.
\bibliography{\filename@base}
\makeatother

>>>>

% User-adder preamble
\usepackage{textcomp} \usepackage{tabu} \usepackage{caption}

\captionsetup{justification = justified} \renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend} \hyphenation{bio-di-ver-si-ty sap-lings re-or-gan-i-za-tion post-dis-tur-bance dis-tur-bance}

\usepackage[left]{lineno} \linenumbers \usepackage{bookmark}

@@

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

\hypertarget{material-and-methods}{%
\section{Material and Methods}\label{material-and-methods}}

\hypertarget{study-site}{%
\subsection{Study site}\label{study-site}}

@@

\begin{table*}[t]

\caption{\label{tab:Tab1}Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou. Treatment intensities are defined by the minimum logging DBH (Diameter at Breast Height), the type of logged species (commercially exploited or not), the density of logged trees, and the total AGB (Above Ground Biomass) lost after treatment.}
\centering
\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\toprule
Treatment & Timber & Thinning & Fuelwood & \%AGB lost\\
\midrule
Control & - & - & - & 0\\
T1, low & DBH $\geq$ 50 cm, commercially exploited species, $\approx$ 10   $trees.ha^{-1}$ & - & - & $[12-33]$\\
T2, intermediate & DBH $\geq$ 50 cm, commercially exploited species, $\approx$ 10  $trees.ha^{-1}$ & DBH $\geq$ 40 cm, non-commercially exploited species, $\approx$ 30   $trees.ha^{-1}$ & - & $[33-56]$\\
T3, high & DBH $\geq$ 50 cm, commercially exploited species, $\approx$ 10  $trees.ha^{-1}$ & DBH $\geq$ 50 cm, non-commercially exploited species, $\approx$ 15  $trees.ha^{-1}$ & 40 cm $\leq$ DBH $\leq$ 50 cm, non-commercially exploited species,\ $\approx$ 15 $trees.ha^{-1}$ & $[35-56]$\\
\bottomrule
\end{tabu}
\end{table*}\ignorespacesafterend

\hypertarget{inventories-protocol-and-dataset-collection}{%
\subsection{Inventories protocol and dataset collection}\label{inventories-protocol-and-dataset-collection}}

@@

\hypertarget{composition-and-diversity-metrics}{%
\subsection{Composition and diversity metrics}\label{composition-and-diversity-metrics}}

@@

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{community-composition}{%
\subsection{Community Composition}\label{community-composition}}

@@

\hypertarget{community-taxonomic-and-functional-diversity}{%
\subsection{Community taxonomic and functional diversity}\label{community-taxonomic-and-functional-diversity}}

@@

\hypertarget{functional-redundancy}{%
\subsection{Functional redundancy}\label{functional-redundancy}}

@@

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

@@

\hypertarget{decoupled-taxonomic-and-functional-trajectories}{%
\subsection{Decoupled taxonomic and functional trajectories}\label{decoupled-taxonomic-and-functional-trajectories}}

@@

\color{red}

We must temper these results with the low number of plots (3) for each treatment. However, the size of each plot is 6.25ha that has been regularly censused over 25 years so that the trajectories have been drawn using 75ha of permanent plot information, a spatial hold very rarely achieved in such studies. Given the relative homogeneity of environmental characteristics of terra firme forests over the Guiana Shield and even the amazonian basin, it seems reasonable that similar trajectories would apply to neighboring areas \citep{Guitet2015, Guitet2018}.
\color{black}

\hypertarget{the-scope-of-the-intermediate-disturbance-hypothesis}{%
\subsection{The scope of the intermediate disturbance hypothesis}\label{the-scope-of-the-intermediate-disturbance-hypothesis}}

Trajectories of taxonomic richness and evenness differed markedly below and above an intensity threshold (Fig. \ref{fig:DivTaxo}).
For low and intermediate disturbance, both taxonomic richness and evenness increased according to the disturbance intensity \citep{Martin2015, Chaudhary2016}.
This suggested that the recruitment of pioneers, previously infrequent or absent, increased the taxonomic richness, and that trees surviving after disturbance remained numerous enough to maintain the richness of the pre-disturbance community \citep{Bongers2009}.
The pioneers thus recruited became abundant, which balanced the usual hyper-dominance of tropical forests and increased the taxonomic evenness \citep{Baraloto2012a}.
Above the intensity threshold, like for treatment 3 (high intensity), the taxonomic richness did not exceed the initial value in the first years following disturbance.
No increase of the taxonomic richness suggested that the richness of surviving trees was lower than this of the pre-disturbance community, and that this difference was not offset by the recruitment of pioneers.
In the Guiana shield, the pool of \color{red}hard \color{black} pioneers specifically recruited after disturbance is restricted to a few common genera (e.g.~\emph{Cecropia spp.}, \emph{Miconia spp.}, \emph{Tapirira spp.}) \citep{Guitet2018}.
\color{red}The disturbance intensity showed a humped-shaped relationship with the post-disturbance increase in taxonomic richness at all the different times after disturbance, and with the increase in taxonomic evenness 10 years after disturbance (Fig. \ref{fig:IDHplot}).\color{black}
Taxonomic richness, and to some extend taxonomic evenness, were maximized at an intermediate intensity, around 20-25\% of AGB lost.

\hypertarget{the-functional-redundancy-explaining-the-taxonomic-functional-decoupling}{%
\subsection{The functional redundancy explaining the taxonomic-functional decoupling}\label{the-functional-redundancy-explaining-the-taxonomic-functional-decoupling}}

@@

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

@@

\hypertarget{acknowledgement}{%
\section{Acknowledgement}\label{acknowledgement}}

@@

\hypertarget{authors-contributions}{%
\section{Author's contributions}\label{authors-contributions}}

@@

\hypertarget{data-availability}{%
\section{Data availability}\label{data-availability}}

@@

\bibliographystyle{mee}
\makeatletter
% The filename has .bib extension the must be eliminated
\filename@parse{references.bib}
% parse stores the file name in base. Extension starts at the first dot, so don't use dots in file names.
\bibliography{\filename@base}
\makeatother
